Do More Subdivisions in Wards Increase Voter Turnout? An Analysis of the 2022 Toronto Municipal Election

### 2023-06-03 outputs/paper/toronto_elections_paper.qmd
Added ggrepel package to library, updated description of packages used, updated code for figure 3 to remove overlapping text, and fixed typo in discussion section.

## Load packages
library(tidyverse)
library(knitr)
library(kableExtra)
library(here)
library(ggplot2)
library(ggbeeswarm)

@@

## Data
The data utilized throughout this paper was obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto]. The data set used is entitled 'Elections - Voter Statistics' [@VoterStatistics]. Data was collected and analyzed using the statistical programming software R [@r], with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `KableExtra` [@rKableExtra], `knitr` [@rknitr], `ggbeeswarm` [@rggbeeswarm], and `here` [@rHere]. A further discussion of the data collection, cleaning, and analysis process can be found later on in this paper.

@@

#Voter turnout by ward
#Code referenced from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.pdf
# & https://stackoverflow.com/questions/66652330/error-ggplot-continuous-x-aesthetic-did-you-forget-aesgroup
finalized_voter_statistics |>
  ggplot(aes(x = factor(turnout), y = n, label = ward)) +
  geom_point(aes(group = turnout)) + 
  geom_beeswarm() +
  geom_smooth(method="lm")+
  geom_text(hjust=-05, vjust=0.5) + 
labs(
    title = "Number of Subdivisions vs Voter Turnout by Ward",
    x = "Turnout (%)",
    y = "Subdivisions",
    label = "Ward"
  ) 
```

@@

Futhermore, Ward 7 (Humber River-Black Creek) had a 22% voter turnout and 47 subdivisions, emphasizing that having fewer subdivisions potentially contributed to having the lowest voter turnout across the city. Ward 10 (Spadina-Fork York), which had the second lowest voter turnout at 24%, accounts for the highest number of subdivisions within a ward at 94. Previous research reveals that younger, more transient communities reside in the Spadina-Fork York riding and are less likely to engage with municipal politics, potentially contributing to the low voter turnout despite the high number of subdivisions [@Turnout22].

@@

There are several limitations with conducting this analysis. The creation of subdivision boundaries within wards lacks transparency and clarity, while impacting the local democratic process. The City Clerk is responsible for drawing up all subdivisions, however the data and information they use to make these decisions is not public and it is challenging to account for any bias introduced during this process. Toronto Elections @ElxnAccessibility clearly outlines their process for finding and establishing accessible polling locations within subdivisions but does not elaborate on the process of establishing subdivisions themselves.

>>>>

## Load packages
library(tidyverse)
library(knitr)
library(kableExtra)
library(here)
library(ggplot2)
library(ggbeeswarm)
library(ggrepel)

@@

## Data
The data utilized throughout this paper was obtained through the City of Toronto’s OpenDataToronto Library [@rOpenDataToronto]. The data set used is entitled 'Elections - Voter Statistics' [@VoterStatistics]. Data was collected and analyzed using the statistical programming software R [@r], with additional support from `tidyverse` [@rTidyverse], `ggplot2` [@rGgplot2], `dplyr` [@rDplyr], `readr` [@rReadr], `tibble` [@rTibble], `janitor` [@rJanitor], `KableExtra` [@rKableExtra], `knitr` [@rknitr], `ggbeeswarm` [@rggbeeswarm], `ggrepel` [@rggrepel], and `here` [@rHere]. A further discussion of the data collection, cleaning, and analysis process can be found later on in this paper.

@@

#Voter turnout by ward
#Code referenced from: https://github.com/christina-wei/INF3104-1-Covid-Clinics/blob/main/outputs/paper/covid_clinics.pdf
# & https://stackoverflow.com/questions/66652330/error-ggplot-continuous-x-aesthetic-did-you-forget-aesgroup
finalized_voter_statistics |>
  ggplot(aes(x = turnout, y = n, label = ward)) +
  # ggplot(aes(x = factor(turnout), y = n, label = ward)) +
  geom_point() + 
  # geom_point(aes(group = turnout)) + 
  # geom_beeswarm() +
  # geom_smooth(method="lm")+
  # geom_text(hjust=-05, vjust=0.5) +
  ggrepel::geom_text_repel() + 
labs(
    title = "Number of Subdivisions vs Voter Turnout by Ward",
    x = "Turnout (%)",
    y = "Number of subdivisions",
    label = "Ward"
  ) +
  theme_classic()
```

@@

Furthermore, Ward 7 (Humber River-Black Creek) had a 22% voter turnout and 47 subdivisions, emphasizing that having fewer subdivisions potentially contributed to having the lowest voter turnout across the city. Ward 10 (Spadina-Fork York), which had the second lowest voter turnout at 24%, accounts for the highest number of subdivisions within a ward at 94. Previous research reveals that younger, more transient communities reside in the Spadina-Fork York riding and are less likely to engage with municipal politics, potentially contributing to the low voter turnout despite the high number of subdivisions [@Turnout22].

@@

There are several limitations with conducting this analysis. The creation of subdivision boundaries within wards lacks transparency and clarity, while impacting the local democratic process. The City Clerk is responsible for drawing up all subdivisions, however the data and information they use to make these decisions is not public and it is challenging to account for any bias introduced during this process. Toronto Elections [@ElxnAccessibility] clearly outlines their process for finding and establishing accessible polling locations within subdivisions but does not elaborate on the process of establishing subdivisions themselves.
